---
title: "Homework 3"
output: html_document
date: "2025-03-25"
---

```{r}
str(wine)
```
## 4898*12
```{r}
summary(wine)
```

```{r}
sum(is.na(wine))
```

```{r}
wine_clean <- na.omit(wine)
```

```{r}
str(wine_clean)
```
## 4870 * 12

# make colnames format better (w/o space)
```{r}
names(wine_clean) <- make.names(names(wine_clean))
```

```{r}
library(psych)
library(corrplot)
```

```{r}
pairs.panels(wine_clean,
             digits = 2,
             pch = 1,
             lm = TRUE, 
             ellipses = FALSE)
```
## 2b: alchohol

```{r}
cor_matrix <- cor(wine_clean, use = "pairwise.complete.obs")
```
```{r}
cor_matrix
```

```{r}
corrplot(cor_matrix, 
         method = "color",
         order = "hclust", 
         addCoef.col = "black", 
         number.cex = 0.7, 
         tl.col = "black", 
         tl.srt = 45)    

```
## 2b: alcohol

```{r}
library(ggplot2)
```
```{r}
ggplot(wine_clean) + aes(x=quality, y = density) + geom_point()  + geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "blue")
```
```{r}
ggplot(wine_clean,
       aes(x=quality, y = alcohol)) +
  geom_point()+
  geom_smooth(method = "lm",
              se = FALSE,
              linetype = "dashed",
              color = "blue")
```



```{r}
lm_wine <-lm(quality ~.,
             data = wine_clean)
```

```{r}
summary(lm_wine)
```
#### fixed acidity, volatile acidity, residual sugar, free sulfur dioxide, density, pH, sulphates, alcohol are significant.

#VIF:
```{r}
library(car)
```

```{r}
vif_wine <- vif(lm_wine)
sort(vif_wine, decreasing = TRUE)
```
## 3b: density >10, residual sugar >10.
## 3b : higher than 10 indicates the multicollinearity (correlation with other variables)

```{r}
lm_wine_w_sig <-lm(quality ~ fixed.acidity + volatile.acidity + residual.sugar + free.sulfur.dioxide + density + pH + sulphates + alcohol,
                   data = wine_clean)
```

```{r}
summary(lm_wine_w_sig)
```
## Neural Network
```{r}
library(neuralnet)
```

# seperate wine_dataset to training data(70%) and testing data(30%)
```{r}
set.seed(42)
wine_index <- sample(nrow(wine_clean),
                     0.7 * nrow(wine_clean),
                     replace = FALSE)

wine_train <- wine_clean[wine_index,]
wine_test <- wine_clean[-wine_index,]
```

```{r}
tail(wine_clean)
```

```{r}
tail(wine_train)
```

```{r}
set.seed(42)
wine_net <- neuralnet(quality ~.,
                      data = wine_train,
                      hidden = 1,
                      lifesign = "minimal",
                      linear.output = TRUE,
                      threshold = 0.01)
```

```{r}
wine_net$result.matrix
```

```{r}
plot(wine_net)
```


```{r}
set.seed(42)
wine_net2 <- neuralnet(quality ~.,
                      data = wine_train,
                      hidden = 2,
                      lifesign = "minimal",
                      linear.output = TRUE,
                      threshold = 0.01)
```
### seems like we need to do normalization
[hidden: 2    thresh: 0.01    rep: 1/1    steps: stepmax	min thresh: 0.347233350612591
Warning: Algorithm did not converge in 1 of 1 repetition(s) within the stepmax.]

```{r}
#functions definition
normalize <- function(x){return((x-min(x))/(max(x)-min(x)))}
denormalize <- function(y,x){return(y*(max(x)-min(x))+min(x))}
```

## normalized dataset: wine_norm
```{r}
wine_norm <- as.data.frame(lapply(wine_clean, normalize))
```

```{r}
head(wine_norm)
```

```{r}
library(dplyr)
wine_norm %>% group_by(quality) %>% summarize( count = n())
```


# seperate normalized wine dataset to traning and testing datasets.
```{r}
set.seed(42)
wine_index <- sample(nrow(wine_norm),
                     0.7 * nrow(wine_norm),
                     replace = FALSE)

wine_train <- wine_norm[wine_index,]
wine_test <- wine_norm[-wine_index,]
```


# node = 1
```{r}
set.seed(42)
wine_net <- neuralnet(quality ~.,
                      data = wine_train,
                      hidden = 1,
                      lifesign = "minimal",
                      linear.output = TRUE,
                      threshold = 0.05)
```

```{r}
wine_net$result.matrix
```

```{r}
plot(wine_net)
```
```{r}
wine_test_wo_quality <- wine_test[, -ncol(wine_test)]
```

```{r}
wine_net.results <- neuralnet::compute(wine_net, wine_test_wo_quality)
```

```{r}
pred_quality_norm <- denormalize(wine_net.results$net.result, wine_norm$quality)
```

```{r}
actual_quality_denorm <- denormalize(wine_test$quality, wine_norm$quality)
```

```{r}
node1_results <- data.frame(actual = actual_quality_denorm,
                      prediction = pred_quality_norm)
```

```{r}
cor(node1_results$actual, node1_results$prediction)
```
## Accuracy: around 51%


# nodes = 2
```{r}
set.seed(42)
wine_net2 <- neuralnet(quality ~.,
                      data = wine_train,
                      hidden = 2,
                      lifesign = "minimal",
                      linear.output = TRUE,
                      threshold = 0.05)
```
```{r}
wine_net2$result.matrix
```

```{r}
plot(wine_net2)
```
```{r}
wine_net2.results <- neuralnet::compute(wine_net2, wine_test_wo_quality)
```

```{r}
pred2_quality_norm <- denormalize(wine_net2.results$net.result, wine_norm$quality)
```

```{r}
node2_results <- data.frame(actual = actual_quality_denorm,
                      prediction = pred2_quality_norm)
```

```{r}
cor(node2_results$actual, node2_results$prediction)
```
# Accuracy: around 54%


# nodes = 3
```{r}
set.seed(42)
wine_net3 <- neuralnet(quality ~.,
                      data = wine_train,
                      hidden = 3,
                      lifesign = "minimal",
                      linear.output = TRUE,
                      threshold = 0.05)
```

```{r}
wine_net3$result.matrix
```

```{r}
plot(wine_net3)
```
```{r}
wine_net3.results <- neuralnet::compute(wine_net3, wine_test_wo_quality)
```

```{r}
pred3_quality_norm <- denormalize(wine_net3.results$net.result, wine_norm$quality)
```


```{r}
node3_results <- data.frame(actual = actual_quality_denorm,
                      prediction = pred3_quality_norm)
```

```{r}
cor(node3_results$actual, node3_results$prediction)
```
# Accuracy: around 55%


# nodes = 4
```{r}
set.seed(42)
wine_net4 <- neuralnet(quality ~.,
                      data = wine_train,
                      hidden = 4,
                      lifesign = "minimal",
                      linear.output = TRUE,
                      threshold = 0.05)
```
```{r}
wine_net4$result.matrix
```

```{r}
plot(wine_net4)
```
```{r}
wine_net4.results <- neuralnet::compute(wine_net4, wine_test_wo_quality)
```

```{r}
pred4_quality_norm <- denormalize(wine_net4.results$net.result, wine_norm$quality)
```


```{r}
node4_results <- data.frame(actual = actual_quality_denorm,
                      prediction = pred4_quality_norm)
```

```{r}
cor(node4_results$actual, node4_results$prediction)
```
# Accuracy: around 56%

# nodes = 5
```{r}
set.seed(42)
wine_net5 <- neuralnet(quality ~.,
                      data = wine_train,
                      hidden = 5,
                      lifesign = "minimal",
                      linear.output = TRUE,
                      threshold = 0.05)
```
```{r}
wine_net5$result.matrix
```

```{r}
plot(wine_net5)
```

```{r}
wine_net5.results <- neuralnet::compute(wine_net5, wine_test_wo_quality)
```

```{r}
pred5_quality_norm <- denormalize(wine_net5.results$net.result, wine_norm$quality)
```

```{r}
node5_results <- data.frame(actual = actual_quality_denorm,
                      prediction = pred5_quality_norm)
```

```{r}
cor(node5_results$actual, node5_results$prediction)
```
# Accuracy: around 57%


#node = 1 with significant variables
```{r}
set.seed(42)
wine_net_w_sig <- neuralnet(quality ~ fixed.acidity + volatile.acidity + residual.sugar + free.sulfur.dioxide + density + pH + sulphates + alcohol,
                      data = wine_train,
                      hidden = 1,
                      lifesign = "minimal",
                      linear.output = TRUE,
                      threshold = 0.05)
```

```{r}
wine_net_w_sig$result.matrix
```

```{r}
plot(wine_net_w_sig)
```
```{r}
wine_net_w_sig.results <- neuralnet::compute(wine_net_w_sig, wine_test_wo_quality)
```

```{r}
pred_w_sig_quality_norm <- denormalize(wine_net_w_sig.results$net.result, wine_norm$quality)
```

```{r}
node1_w_sig_results <- data.frame(actual = actual_quality_denorm,
                      prediction = pred_w_sig_quality_norm)
```

```{r}
cor(node1_w_sig_results$actual, node1_w_sig_results$prediction)
```
#node = 2 with significant variables
```{r}
set.seed(42)
wine_net2_w_sig <- neuralnet(quality ~ fixed.acidity + volatile.acidity + residual.sugar + free.sulfur.dioxide + density + pH + sulphates + alcohol,
                      data = wine_train,
                      hidden = 2,
                      lifesign = "minimal",
                      linear.output = TRUE,
                      threshold = 0.05)
```

```{r}
wine_net2_w_sig$result.matrix
```

```{r}
plot(wine_net2_w_sig)
```
```{r}
wine_net2_w_sig.results <- neuralnet::compute(wine_net2_w_sig, wine_test_wo_quality)
```

```{r}
pred2_w_sig_quality_norm <- denormalize(wine_net2_w_sig.results$net.result, wine_norm$quality)
```

```{r}
node2_w_sig_results <- data.frame(actual = actual_quality_denorm,
                      prediction = pred2_w_sig_quality_norm)
```

```{r}
cor(node2_w_sig_results$actual, node2_w_sig_results$prediction)
```
# Accuracy: around 54.04%


#node = 3 with significant variables
```{r}
set.seed(42)
wine_net3_w_sig <- neuralnet(quality ~ fixed.acidity + volatile.acidity + residual.sugar + free.sulfur.dioxide + density + pH + sulphates + alcohol,
                      data = wine_train,
                      hidden = 3,
                      lifesign = "minimal",
                      linear.output = TRUE,
                      threshold = 0.05)
```

```{r}
wine_net3_w_sig$result.matrix
```

```{r}
plot(wine_net3_w_sig)
```
```{r}
wine_net3_w_sig.results <- neuralnet::compute(wine_net3_w_sig, wine_test_wo_quality)
```

```{r}
pred3_w_sig_quality_norm <- denormalize(wine_net3_w_sig.results$net.result, wine_norm$quality)
```

```{r}
node3_w_sig_results <- data.frame(actual = actual_quality_denorm,
                      prediction = pred3_w_sig_quality_norm)
```

```{r}
cor(node3_w_sig_results$actual, node3_w_sig_results$prediction)
```
# Accuracy: around 54.3%

#node = 4 with significant variables
```{r}
set.seed(42)
wine_net4_w_sig <- neuralnet(quality ~ fixed.acidity + volatile.acidity + residual.sugar + free.sulfur.dioxide + density + pH + sulphates + alcohol,
                      data = wine_train,
                      hidden = 4,
                      lifesign = "minimal",
                      linear.output = TRUE,
                      threshold = 0.05)
```

```{r}
wine_net4_w_sig$result.matrix
```

```{r}
plot(wine_net4_w_sig)
```
```{r}
wine_net4_w_sig.results <- neuralnet::compute(wine_net4_w_sig, wine_test_wo_quality)
```

```{r}
pred4_w_sig_quality_norm <- denormalize(wine_net4_w_sig.results$net.result, wine_norm$quality)
```

```{r}
node4_w_sig_results <- data.frame(actual = actual_quality_denorm,
                      prediction = pred4_w_sig_quality_norm)
```

```{r}
cor(node4_w_sig_results$actual, node4_w_sig_results$prediction)
```

# Accuracy: around 56%

#node = 5 with significant variables
```{r}
set.seed(42)
wine_net5_w_sig <- neuralnet(quality ~ fixed.acidity + volatile.acidity + residual.sugar + free.sulfur.dioxide + density + pH + sulphates + alcohol,
                      data = wine_train,
                      hidden = 5,
                      lifesign = "minimal",
                      linear.output = TRUE,
                      threshold = 0.05)
```

```{r}
wine_net5_w_sig$result.matrix
```

```{r}
plot(wine_net5_w_sig)
```
```{r}
wine_net5_w_sig.results <- neuralnet::compute(wine_net5_w_sig, wine_test_wo_quality)
```

```{r}
pred5_w_sig_quality_norm <- denormalize(wine_net5_w_sig.results$net.result, wine_norm$quality)
```

```{r}
node5_w_sig_results <- data.frame(actual = actual_quality_denorm,
                      prediction = pred5_w_sig_quality_norm)
```

```{r}
cor(node5_w_sig_results$actual, node5_w_sig_results$prediction)
```
# Accuracy: around 56%