---
title: "Heart Analysis"
output: html_document
date: "2025-02-02"
---

```{r}
library(dplyr)
library(ggplot2)
```

```{r}
summary(HeartFailure)
```

```{r}
ggplot(HeartFailure, aes( x = ejection_fraction, y = age)) + 
  geom_point(aes(colour = DEATH_EVENT, shape = DEATH_EVENT), size = 2) +
  xlab("Ejection Fraction") + ylab("Age") + ggtitle("Impact of Ejection Fraction and Age on Death Events") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(HeartFailure, aes( x = ejection_fraction, y = serum_creatinine  )) + 
  geom_point(aes(colour = DeathEvent, shape = DeathEvent), size = 2) +
  xlab("Ejection Fraction") + ylab("Serum Creatinine") + ggtitle("Impact of Ejection Fraction and Serum Creatinine on Death Events") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
Heart_Clean <- HeartFailure[c("age", 
                               "creatinine_phosphokinase", 
                               "ejection_fraction", 
                               "platelets", 
                               "serum_creatinine", 
                               "serum_sodium", 
                               "time",
                               "DEATH_EVENT")]
```


```{r}
Heart_Clean$DeathEvent <- lapply(Heart_Clean$DEATH_EVENT, function(x){
  if (x == "0")
    Heart_Clean$DEATH_EVENT <- -1
  else if(x == "1")
    Heart_Clean$DEATH_EVENT <- 1
  else
    Heart_Clean$DEATH_EVENT <- NULL
})
```

```{r}
Heart_Subset <- Heart_Clean[c("age", 
                               "creatinine_phosphokinase", 
                               "ejection_fraction", 
                               "platelets", 
                               "serum_creatinine", 
                               "serum_sodium", 
                               "time",
                               "DeathEvent")]
```


```{r}
perceptron <- function(X, y, numEpochs) {
  results <- list()
  w <- runif(ncol(X), -10, 10) #Initalize weights
  # For loop - number of generations(epochs) - number of times dataset is ran through
  for(j in 1:numEpochs) {
  predictedResult <- numeric(length=100) # Initalize predictedResult vector
  numIncorrect = 0 # Keeps track of # of missclassified points
  # For loop - loop throught dataset
  for(i in 1:length(y)) {
  xi = as.numeric(unlist(X[i,])) # Convert dataframe to vector
  predictedResult[i] = sign(w %*% xi) # Predict the point
  # If predicted point is incorrect - change weight
  if(predictedResult[i] != y[i]) {
  numIncorrect = numIncorrect + 1 # Add one to # of missclassified points
  w <- w + as.numeric(y[i]) * xi # Update the weight w <- w + WiXi
  }
}
# Print results of this generation(epoch)
cat("\nEpoch #: ", j)
cat("\nNumber Incorrect: ", numIncorrect)
cat("\nFinal Weight: ", w)
  }
}
```


```{r}
set.seed(123)

Heart_index <- sample(nrow(Heart_Subset), 0.7 * nrow(Heart_Subset))
```

```{r}
Heart_train <- Heart_Subset[Heart_index,]
```

```{r}
Heart_test <- Heart_Subset[-Heart_index,]
```

```{r}
X <- Heart_train[, c("ejection_fraction", "age")] # Input Matrix
```

```{r}
Y <- Heart_train$DeathEvent
```


=== X1: "ejection_fraction", "age" ====


```{r}
set.seed(123)
perceptron(X, Y, 25)
```

Epoch #:  22
# Pick the weight set that results in the lowest incorrect number count : 68
# Weight: -161.2484 97.7661
# "ejection_fraction" "age"

```{r}
W1 = c(-161.2484, 97.7661)
```

```{r}
W1_Heart_test <- Heart_test
```

```{r}
W1_Heart_test$W1_ejection_fraction <- W1_Heart_test$ejection_fraction * W1[1]
W1_Heart_test$W1_age <- W1_Heart_test$age * W1[2]

```

```{r}
W1_Heart_test$predicit <- rowSums(W1_Heart_test [,c("W1_ejection_fraction", "W1_age")])
```

```{r}
w1_perceptron_predict_table <- table(W1_Heart_test$DeathEvent == 1, W1_Heart_test$predicit >0) + table(W1_Heart_test$DeathEvent == -1, W1_Heart_test$predicit < 0)
w1_perceptron_predict_table
```


```{r}
sum(diag(w1_perceptron_predict_table)/sum(w1_perceptron_predict_table))
```
# Perceptron Predict of "ejection_fraction" & "age" on Death Events
# Accuracy : 0.633



=== X2: "ejection_fraction", "serum_creatinine" ===


```{r}
X2 <- Heart_train[, c("ejection_fraction", "serum_creatinine")] 
```


```{r}
set.seed(123)

perceptron(X2, Y, 25)

```

Epoch #:  15
# Pick the weight set that results in the lowest incorrect number count : 67
# Weight:  -8.24845 316.1261
# "ejection_fraction" "serum_creatinine"


```{r}
W2 = c(-8.24845, 316.1261)
```


```{r}
W2_Heart_test <- Heart_test
```


```{r}
W2_Heart_test$W2_ejection_fraction <- W2_Heart_test$ejection_fraction * W2[1]
W2_Heart_test$W2_serum_creatinine <- W2_Heart_test$serum_creatinine * W2[2]

```


```{r}
W2_Heart_test$predicit <- rowSums(W2_Heart_test [,c("W2_ejection_fraction", "W2_serum_creatinine")])
```


```{r}
w2_perceptron_predict_table <- table(W2_Heart_test$DeathEvent == 1, W2_Heart_test$predicit >0) + table(W2_Heart_test$DeathEvent == -1, W2_Heart_test$predicit < 0)
w2_perceptron_predict_table
```


```{r}
sum(diag(w2_perceptron_predict_table)/sum(w2_perceptron_predict_table))
```
# Perceptron Predict of "ejection_fraction" & "serum_creatinine" on Death Events
# Accuracy : 0.644



=== X3: "ejection_fraction", "serum_creatinine", "age" ===


```{r}
X3 <- Heart_train[, c("ejection_fraction", "serum_creatinine", "age")] # Input Matrix

```


```{r}
set.seed(123)

perceptron(X3, Y, 25)
```

Epoch #:  25
Number Incorrect:  66
Final Weight:  -166.2484 503.5461 83.17954

```{r}
W3 = c(-166.2484, 503.5461, 83.17954)
```


```{r}
W3_Heart_test <- Heart_test
```


```{r}
W3_Heart_test$W3_ejection_fraction <- W3_Heart_test$ejection_fraction * W3[1]
W3_Heart_test$W3_serum_creatinine <- W3_Heart_test$serum_creatinine * W3[2]
W3_Heart_test$W3_age <- W3_Heart_test$age * W3[3]
```


```{r}
W3_Heart_test$predicit <- rowSums(W3_Heart_test [,c("W3_ejection_fraction", "W3_serum_creatinine", "W3_age")])
```


```{r}
w3_perceptron_predict_table <- table(W3_Heart_test$DeathEvent == 1, W3_Heart_test$predicit >0) + table(W3_Heart_test$DeathEvent == -1, W3_Heart_test$predicit < 0)
w3_perceptron_predict_table
```

```{r}
sum(diag(w3_perceptron_predict_table)/sum(w3_perceptron_predict_table))
```
# Perceptron Predict of "ejection_fraction" & "serum_creatinine" & "age"" on Death Events
# Accuracy : 0.6667



=== X4: "age", "ejection_fraction", "serum_creatinine", "serum_sodium" ===

```{r}
X4 <- Heart_train[, c("age", 
                      "ejection_fraction",
                      "serum_creatinine", 
                      "serum_sodium")
]

```


```{r}
set.seed(123)

perceptron(X4, Y, 25)
```


Epoch #:  17
# Pick the weight set that results in the lowest incorrect number count : 62
# Weight: 523.7516 -730.2339 425.0195 -72.33965
# "age", "ejection_fraction",  "serum_creatinine", "serum_sodium"

```{r}
W4 = c(523.7516, -730.2339, 425.0195, -72.33965)
```


```{r}
W4_Heart_test <- Heart_test
```

```{r}

W4_Heart_test$W4_age <- W4_Heart_test$age * W4[1]
W4_Heart_test$W4_ejection_fraction <- W4_Heart_test$ejection_fraction * W4[2]
W4_Heart_test$W4_serum_creatinine <- W4_Heart_test$serum_creatinine * W4[3]
W4_Heart_test$W4_serum_sodium <- W4_Heart_test$serum_sodium * W4[4]

```


```{r}
W4_Heart_test$predicit <- rowSums(W4_Heart_test [,c("W4_age", "W4_ejection_fraction", "W4_serum_creatinine", "W4_serum_sodium")])
```


```{r}
w4_perceptron_predict_table <- table(W4_Heart_test$DeathEvent == 1, W4_Heart_test$predicit >0) + table(W4_Heart_test$DeathEvent == -1, W4_Heart_test$predicit < 0)
w4_perceptron_predict_table
```

```{r}
sum(diag(w4_perceptron_predict_table)/sum(w4_perceptron_predict_table))
```
# Perceptron Predict of "age" & "ejection_fraction" & "serum_creatinine" & "serum_sodium" on Death Events
# Accuracy : 0.7111




####### SVM


```{r}
library(e1071)
```

```{r}
HeartFailure_SVM1 <- HeartFailure[, c("DEATH_EVENT", "age", "smoking")]
```

```{r}
HeartFailure_SVM1$DEATH_EVENT <- as.factor(HeartFailure_SVM1$DEATH_EVENT)
```


3a:
#### age, smoking

```{r}
set.seed(123)

heart_svm1_index <- sample(nrow(HeartFailure_SVM1), 0.7 * nrow(HeartFailure_SVM1), replace = FALSE)
```

```{r}
heart_svm1_train <- HeartFailure_SVM1[heart_svm1_index,]
heart_svm1_test <- HeartFailure_SVM1[-heart_svm1_index,]
```

```{r}
set.seed(123)
heart_svm1_model <-svm(DEATH_EVENT ~ age + smoking, data = heart_svm1_train)
```

```{r}
print(heart_svm1_model)
```

```{r}
summary(heart_svm1_model)

```

```{r}
plot(heart_svm1_model, heart_svm1_train, age ~ smoking)
```

```{r}
heart_svm1_predict <- predict(heart_svm1_model, heart_svm1_test[,-1], type = "response")
```

```{r}
svm1_predicttable <- table(heart_svm1_test$DEATH_EVENT == '1', heart_svm1_predict == '1')
svm1_predicttable
```

```{r}
sum(diag(svm1_predicttable))/sum(svm1_predicttable)
```
# accuracy : 0.6889


#### 3b
### "diabetes", "creatinine_phosphokinase"

```{r}
HeartFailure_SVM2 <- HeartFailure[, c("DEATH_EVENT", "diabetes", "creatinine_phosphokinase")]
```


```{r}
HeartFailure_SVM2$DEATH_EVENT <- as.factor(HeartFailure_SVM2$DEATH_EVENT)
```

```{r}
set.seed(123)
heart_svm2_index <- sample(nrow(HeartFailure_SVM2), 0.7 * nrow(HeartFailure_SVM2), replace = FALSE)
```

```{r}
heart_svm2_train <- HeartFailure_SVM2[heart_svm2_index,]
heart_svm2_test <- HeartFailure_SVM2[-heart_svm2_index,]
```

```{r}
heart_svm2_model <-svm(DEATH_EVENT ~ diabetes + creatinine_phosphokinase, data = heart_svm2_train)
```

```{r}
print(heart_svm2_model)
```

```{r}
summary(heart_svm2_model)

```

```{r}
plot(heart_svm2_model, heart_svm2_train, diabetes ~ creatinine_phosphokinase)
```
# People who have diabetes and higher Creatinine Phosphokinase(>6500) might have a higher rate of Death than those who dont have diabetes and lower Creatinine Phosphokinase (<6500).

```{r}
heart_svm2_predict <- predict(heart_svm2_model, heart_svm2_test[,-1], type = "response")
```


```{r}
heart_svm2_predict
```

```{r}
svm2_predicttable <- table(heart_svm2_test$DEATH_EVENT == '1', heart_svm2_predict == '1')
svm2_predicttable
```

```{r}
sum(diag(svm2_predicttable))/sum(svm2_predicttable)
```
## accuracy: 0.6778




## 3c:
## combine all 4 above


```{r}
HeartFailure_SVM3 <- HeartFailure[, c("DEATH_EVENT", "age", "smoking", "diabetes", "creatinine_phosphokinase")]
```

```{r}
HeartFailure_SVM3$DEATH_EVENT <- as.factor(HeartFailure_SVM3$DEATH_EVENT)

```


```{r}
set.seed(123)

heart_svm3_index <- sample(nrow(HeartFailure_SVM3), 0.7 * nrow(HeartFailure_SVM3), replace = FALSE)
```

```{r}
heart_svm3_train <- HeartFailure_SVM3[heart_svm3_index,]
heart_svm3_test <- HeartFailure_SVM3[-heart_svm3_index,]
```

```{r}
heart_svm3_model <-svm(DEATH_EVENT ~., data = heart_svm3_train)
```

```{r}
print(heart_svm3_model)
```

```{r}
summary(heart_svm3_model)

```

```{r}
plot(heart_svm3_model, heart_svm3_train, age ~ creatinine_phosphokinase, slice = list(smoking = 1, diabetes = 1))
```
```{r}
plot(heart_svm3_model, heart_svm3_train, age ~ creatinine_phosphokinase, slice = list(smoking = 0, diabetes = 1))

```

###  plots interpretation:
Compare "people who don’t smoke (smoking = 0) and have diabetes (diabetes = 1)" with "people who smoke and have diabetes."

The model predicts that people who smoke tend to die earlier, as shown by the larger RED "Died" region and started around age "80", whereas for non-smokers (smoking = 0), it starts around age 90.
I mainly compare the low creatinine phosphokinase (CPK) range because the data in the high CPK range is sparse, making the model's predictions in that area less reliable.

```{r}
heart_svm3_predict <- predict(heart_svm3_model, heart_svm3_test[,-1], type = "response")
```

```{r}
svm3_predicttable <- table(heart_svm3_test$DEATH_EVENT == '1', heart_svm3_predict == '1')
svm3_predicttable
```

```{r}
sum(diag(svm3_predicttable))/sum(svm3_predicttable)
```
# accuracy: 0.7

## 3d:
## "DeathEvent", "age", "smoking", "diabetes", "sex", "high_blood_pressure", "ejection_fraction"


```{r}
HeartFailure_SVM4 <- HeartFailure[, c("DEATH_EVENT", "age", "smoking", "diabetes", "sex", "high_blood_pressure", "ejection_fraction")]
```

```{r}
HeartFailure_SVM4$DEATH_EVENT <- as.factor(HeartFailure_SVM4$DEATH_EVENT)
```

```{r}
set.seed(123)
heart_svm4_index <- sample(nrow(HeartFailure_SVM4), 0.7 * nrow(HeartFailure_SVM4), replace = FALSE)
```

```{r}
heart_svm4_train <- HeartFailure_SVM4[heart_svm4_index,]
heart_svm4_test <- HeartFailure_SVM4[-heart_svm4_index,]
```

```{r}
heart_svm4_model <-svm(DEATH_EVENT ~., data = heart_svm4_train)
```

```{r}
print(heart_svm4_model)
```

```{r}
summary(heart_svm4_model)

```

```{r}
plot(heart_svm4_model, heart_svm4_train, age ~ ejection_fraction,
     slice = list(smoking = 1, diabetes = 1, sex = 0, high_blood_pressure = 1))
```
↑ smoker/diabetes/femal/high blood pressure


```{r}
plot(heart_svm4_model, heart_svm4_train, age ~ ejection_fraction,
     slice = list(smoking = 1, diabetes = 0, sex = 0, high_blood_pressure = 0))
```
↑ smoker/no diabetes/female/no high blood pressure


## We can see that smokers with diseases have a larger yellow area(Died) than those with no diseases.


```{r}
plot(heart_svm4_model, heart_svm4_train, age ~ ejection_fraction,
     slice = list(smoking = 0, diabetes = 1, sex = 0, high_blood_pressure = 1))

```
↑ not smoker/diabetes/female/high blood pressure

# We can see that even non-smokers with all diseases still have a large red area(Died), unfortunately. However, compared to smokers with all diseases, the red area is smaller.


```{r}
plot(heart_svm4_model, heart_svm4_train, age ~ ejection_fraction,
     slice = list(smoking = 0, diabetes = 0, sex = 0, high_blood_pressure = 0))
```
↑ not smoker/no diabetes/female/no high blood pressure


```{r}
heart_svm4_predict <- predict(heart_svm4_model, heart_svm4_test[,-1], type = "response")
```

```{r}
svm4_predicttable <- table(heart_svm4_test$DEATH_EVENT == '1', heart_svm4_predict == '1')
svm4_predicttable
```

```{r}
sum(diag(svm4_predicttable))/sum(svm4_predicttable)
```
# accuracy : 0.6889

