---
title: "Homework2"
output: html_document
date: "2025-02-25"
---

```{r}
library(corrplot)
library(neuralnet)
library(caret)
```

=== Clean Data ===

# convert Gender to dummy
# Female = 0, Male = 1
```{r}
patient <- indian_liver_patient
```

```{r}
patient$Gender <- ifelse(patient$Gender == "Female", 0,1)
```

```{r}
sum(is.na(patient))
```

```{r}
str(patient)
```
583 * 11

# eliminate NA

```{r}
patient_clean <- na.omit(patient)
```

```{r}
str(patient_clean)
```
579 * 11


## Correlation 
```{r}
library(corrplot)
```


```{r}
cor_matrix <- cor(patient_clean)
```


```{r}
corrplot(round(cor_matrix, 3), 
         method = "color",
         order = "hclust", 
         addCoef.col = "black", 
         number.cex = 0.7, 
         tl.col = "black", 
         tl.srt = 45)    

```
2.b Direct_Bilirubin

# logit 
```{r}
model_logit <- glm(Disease ~. ,data = patient_clean, family = "binomial")
```
```{r}
summary(model_logit)
```
3a.
Age, Alamine_Aminotransferase, Total_Protiens, Albumin

```{r}
model_logit2 <- glm(Disease ~Age + Alamine_Aminotransferase + Total_Protiens + Albumin ,data = patient_clean, family = "binomial")
```

```{r}
summary(model_logit2)
```

# NeuralNet
```{r}
library(neuralnet)
```

```{r}
set.seed(42)
patient_index <- sample(nrow(patient_clean), 0.7 * nrow(patient_clean),
                        replace = FALSE)
patient_train <- patient_clean[patient_index, ]
patient_test <- patient_clean[-patient_index, ]

```

```{r}
tail(patient_clean)
```
```{r}
tail(patient_train)
```
```{r}
tail(patient_test)
```

a: hidden = 1
```{r}
set.seed(42)
model_neural <- neuralnet(Disease ~.,
                          data = patient_train,
                          hidden = 1,
                          lifesign = "minimal",
                          linear.output = FALSE,
                          threshold = 0.01)
```

```{r}
plot(model_neural)
```

```{r}
result_model_neural1 <- compute(model_neural, patient_test[1:10])
```

```{r}
df_result_model_neural1 <- data.frame(actual=patient_test$Disease, prediction=result_model_neural1$net.result)
```

```{r}
df_result_model_neural1$round_predic <- round(df_result_model_neural1$prediction)
```

```{r}
df_result_model_neural1
```

```{r}
df_result_model_neural1$actual <- as.factor(df_result_model_neural1$actual)
df_result_model_neural1$round_predic <- as.factor(df_result_model_neural1$round_predic)
```

```{r}
confusionMatrix(df_result_model_neural1$actual, df_result_model_neural1$round_predic)
```
# predict numbers are all 1 thats why there is an error (need levels)

```{r}
table_neural1_acc = table(df_result_model_neural1$actual, df_result_model_neural1$round_predic)
```

```{r}
table_neural1_acc
```
```{r}
122/174
```

============
# 3a.model1. 
accuracy: 70%
=============


3a. model 2 => hidden nodes = 2
```{r}
set.seed(42)
model_neural2 <- neuralnet(Disease ~.,
                           data = patient_train,
                           hidden = 2,
                           lifesign = "minimal",
                           linear.output = FALSE,
                           threshold = 0.01)
```

```{r}
plot(model_neural2)
```

```{r}
result_model_neural2 <- compute(model_neural2, patient_test[1:10])
```

```{r}
df_result_model_neural2 <- data.frame(actual=patient_test$Disease, prediction=result_model_neural2$net.result)
```


```{r}
df_result_model_neural2$round_predic <- round(df_result_model_neural2$prediction)
```


```{r}
df_result_model_neural2
```
```{r}
df_result_model_neural2$actual <- as.factor(df_result_model_neural2$actual)
df_result_model_neural2$round_predic <- as.factor(df_result_model_neural2$round_predic)
```

```{r}
confusionMatrix(df_result_model_neural2$actual, df_result_model_neural2$round_predic)
```
# Accuracy : 70%


a: hidden = 3
```{r}
set.seed(42)
model_neural3 <- neuralnet(Disease ~.,
                           data = patient_train,
                           hidden = 3 ,
                           lifesign = "minimal",
                           linear.output = FALSE,
                           threshold = 0.01)
```

```{r}
plot(model_neural3)
```

```{r}
result_model_neural3 <- compute(model_neural3, patient_test[1:10])
```

```{r}
df_result_model_neural3 <- data.frame(actual=patient_test$Disease, prediction=result_model_neural3$net.result)
```

```{r}
df_result_model_neural3$round_predic <- round(df_result_model_neural3$prediction)
```


```{r}
df_result_model_neural3
```
```{r}
df_result_model_neural3$actual <- as.factor(df_result_model_neural3$actual)
df_result_model_neural3$round_predic <- as.factor(df_result_model_neural3$round_predic)
```

```{r}
confusionMatrix(df_result_model_neural3$actual, df_result_model_neural3$round_predic)
```
# Accuracy: 71%



# 3b w significant variables 

```{r}
set.seed(42)
model_neural_w_sig <- neuralnet(Disease ~ Age + Alamine_Aminotransferase + Total_Protiens + Albumin,
                          data = patient_train,
                          hidden = 1,
                          lifesign = "minimal",
                          linear.output = FALSE,
                          threshold = 0.01)
```
```{r}
plot(model_neural_w_sig)
```

```{r}
result_model_neural_w_sig <- compute(model_neural_w_sig, patient_test[1:10])
```

```{r}
df_result_model_neural_w_sig <- data.frame(actual=patient_test$Disease, prediction=result_model_neural_w_sig$net.result)
```

```{r}
df_result_model_neural_w_sig$round_predic <- round(df_result_model_neural_w_sig$prediction)
```


```{r}
df_result_model_neural_w_sig
```
```{r}
df_result_model_neural_w_sig$actual <- as.factor(df_result_model_neural_w_sig$actual)
df_result_model_neural_w_sig$round_predic <- as.factor(df_result_model_neural_w_sig$round_predic)
```

```{r}
confusionMatrix(df_result_model_neural3$actual, df_result_model_neural3$round_predic)
```
# accuracy : 0.71

====

```{r}
set.seed(42)
model_neural_w_sig2 <- neuralnet(Disease ~ Age + Alamine_Aminotransferase + Total_Protiens + Albumin,
                          data = patient_train,
                          hidden = 2,
                          lifesign = "minimal",
                          linear.output = FALSE,
                          threshold = 0.01)
```
```{r}
plot(model_neural_w_sig2)
```

```{r}
result_model_neural_w_sig2 <- compute(model_neural_w_sig2, patient_test[1:10])
```

```{r}
df_result_model_neural_w_sig2 <- data.frame(actual=patient_test$Disease, prediction=result_model_neural_w_sig2$net.result)
```

```{r}
df_result_model_neural_w_sig2$round_predic <- round(df_result_model_neural_w_sig2$prediction)
```


```{r}
df_result_model_neural_w_sig2
```
```{r}
df_result_model_neural_w_sig2$actual <- as.factor(df_result_model_neural_w_sig2$actual)
df_result_model_neural_w_sig2$round_predic <- as.factor(df_result_model_neural_w_sig2$round_predic)
```

```{r}
confusionMatrix(df_result_model_neural_w_sig2$actual, df_result_model_neural_w_sig2$round_predic)
```
# predict number are all 1 thats why there is an error

```{r}
table_neural2_w_sig_acc = table(df_result_model_neural_w_sig2$actual, df_result_model_neural_w_sig2$round_predic)
```

```{r}
table_neural2_w_sig_acc
```
```{r}
122/174
```
# accuracy : 70%


```{r}
set.seed(42)
model_neural_w_sig3 <- neuralnet(Disease ~ Age + Alamine_Aminotransferase + Total_Protiens + Albumin,
                          data = patient_train,
                          hidden = 3,
                          lifesign = "minimal",
                          linear.output = FALSE,
                          threshold = 0.01)
```
```{r}
plot(model_neural_w_sig3)
```

```{r}
result_model_neural_w_sig3 <- compute(model_neural_w_sig3, patient_test[1:10])
```

```{r}
df_result_model_neural_w_sig3 <- data.frame(actual=patient_test$Disease, prediction=result_model_neural_w_sig3$net.result)
```

```{r}
df_result_model_neural_w_sig3$round_predic <- round(df_result_model_neural_w_sig3$prediction)
```


```{r}
df_result_model_neural_w_sig3
```
```{r}
df_result_model_neural_w_sig3$actual <- as.factor(df_result_model_neural_w_sig3$actual)
df_result_model_neural_w_sig3$round_predic <- as.factor(df_result_model_neural_w_sig3$round_predic)
```

```{r}
confusionMatrix(df_result_model_neural_w_sig3$actual, df_result_model_neural_w_sig3$round_predic)
```
# accuracy: 74%